---
title: "A2"
author: "Jefferson Li, Arib Shaikh"
date: "15/02/2021"
header-includes:
   - \usepackage{amsthm}
   - \usepackage{accents}
output: pdf_document

---
Set Seed
```{r seed}
set.seed(1005057368) 
```
## Q1
### a)

Note that $$
\hat{\beta_1} \sim N(\beta_1, \sigma^2/Sxx)\\
$$
We know $\beta_1 = 4, \sigma^2 = 25$.

We can calculate Sxx from the following code
```{r}
Xs = c(4, 8, 12, 16, 20)
Sxx = 0
for (x in Xs) {
  Sxx = Sxx + (x - mean(Xs))^2
}
Sxx
```
Therefore 
$$
\hat{\beta_1} \sim N(4, 25/`r Sxx`)\\
$$
We want to calculate,
\begin{align}
    P(|\hat{\beta_1} - \beta_1| >1)
    &= P(|\hat{\beta_1} - 4| >1)\\
    &= P(\hat{\beta_1} - 4 > 1) \quad or \quad P(\hat{\beta_1} - 4 < -1)\\
    &= P(\hat{\beta_1} > 5)+ P(\hat{\beta_1}< 3) && \text{Since they are disjoint}
\end{align}

This probability can be calculated in r with the following code

```{r}
prob = pnorm(3, 4, sqrt(25/Sxx)) + (1-pnorm(5, 4, sqrt(25/Sxx)))
```
Therefore, $$ P(|\hat{\beta_1} - \beta_1| >1) = `r prob`$$ 

### b)



```{r}
Xs = c(4, 8, 12, 16, 20)

errors = rnorm(5,0,5)
errors
```

```{r}
Ys = rep(0, 5)
for (i in 1:length(Ys)) {
    Ys[i] = 20 + 4*Xs[i] + errors[i]
}
Ys
```
```{r}
Sxy = 0
for (i in 1:length(Ys)) {
    Sxy = Sxy + (Xs[i] - mean(Xs))*(Ys[i] - mean(Ys))
}
Sxy
```

```{r}
Bhat1 = Sxy / Sxx
Bhat0 = mean(Ys) - Bhat1*mean(Xs)
Bhat1
Bhat0
```


```{r}
X0 = 10
Yhat0 =  Bhat0 + Bhat1 * X0
Yhat0
```

```{r}
fit = lm(formula = Ys ~ Xs)
CI = predict(fit, data.frame(Xs=10), interval="confidence")
```
The 95% confidence interval for E(Y0) when X0 = 10 is given by $$[`r CI[2]`, `r CI[3]`]$$

### c)

```{r}
Bhat0s = rep(0,1000)
Bhat1s = rep(0,1000)
lowers = rep(0,1000)
uppers = rep(0,1000)

Xs = c(4, 8, 12, 16, 20)

for (i in 1:length(Bhat0s)) {
  errors = rnorm(5,0,5)

  Ys = rep(0, 5)
  for (j in 1:length(Ys)) {
      Ys[j] = 20 + 4*Xs[j] + errors[j]
  }
  fit = lm(formula = Ys ~ Xs)
  
  Bhat0s[i] = fit$coefficients[1]
  Bhat1s[i] = fit$coefficients[2]
  CI = predict(fit, data.frame(Xs=10), interval="confidence")
  lowers[i] = CI[2]
  uppers[i] = CI[3]
 }
```
```{r}
hist(Bhat1s, xlab = "beta 1 hat")
mean(Bhat1s)
sd(Bhat1s)
```
The sample mean of $`r mean(Bhat1s)`$ is consistent with the theoretical mean of $4$.

The sample standard deviation of $`r sd(Bhat1s)`$ is consistent with the theoretical standard deviation of $\sqrt{\sigma^2/Sxx} = \sqrt{25/160} = 0.3952$.

```{r}
hist(Bhat0s, xlab = "beta 0 hat")
mean(Bhat0s)
sd(Bhat0s)
```
The sample mean of $`r mean(Bhat0s)`$ is consistent with the theoretical mean of $20$.

The sample standard deviation of $`r sd(Bhat0s)`$ is consistent with the theoretical standard deviation of $\sqrt{\frac{1}{n}+\frac{\bar{x}^2}{Sxx}}\sigma = \sqrt{\frac{1}{5}+\frac{12^2}{160}}5 = 5.244$.

### d)


```{r}
numsWithinCI = 0

for (i in 1:length(lowers)) {
  Ey = 20 + 4*X0
  if (Ey > lowers[i] && Ey < uppers[i]){
    numsWithinCI = numsWithinCI + 1
  }
}
numsWithinCI
```
This result of 951/1000 is consistent with the theoretical proportion of 950/1000


## Q4
### a)
### 1)
```{r}
Xs = c(3,5,4,6,7)
Ys = c(4,6.5,5,7,7.5)
fit = lm(formula = Ys ~ Xs,x=TRUE)
designMatrix = matrix(model.matrix(fit), nrow = 5, ncol = 2, byrow = FALSE)
designMatrix 
```

### 2)

```{r}
coeffs = solve(t(designMatrix)%*%designMatrix)%*%t(designMatrix)%*%Ys
coeffs
```

### 3)
```{r}
varCovarMatrix = solve(t(designMatrix)%*%designMatrix)
varCovarMatrix
```


### b)

### 1)
```{r}
varCovarMatrix[1,2]
```

### 2)
```{r}
varCovarMatrix[1,1]
```

### 3)
```{r}
varCovarMatrix[2,2]
```

### c)

```{r}
H = designMatrix%*%solve(t(designMatrix)%*%designMatrix)%*%t(designMatrix)
H
```

### d)

```{r}
sigma2 = 4
VarE = (1 - H)*sigma2
VarE
```



